---
title: "Weather and Corn Yield Regressions"
author: "Kaydee Barker"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab) #pull in matlab files
library(rnassqs) #access to USDA national agriculture data
```

## Weather Data Analysis

### Load the PRISM daily maximum temperatures

```{r tmax data}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1] #first county, all days, first year
t_1981_c1[366] #check for leap year (366 days)
plot(1:366, t_1981_c1, type = "l") #base r plot

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```

```{r tidying up}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years) #add dimension names

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax") #name columns
tmaxdf <- tibble(tmaxdf) #tidyverse table

```

## Temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends}

tmaxdf$doy <- as.numeric(tmaxdf$doy) 
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>% #day 152= June 1, 243= Aug 31
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax) #not sig

```

### Winter Temperatures - Winneshiek County

```{r winter temps}

winnewinter <- tmaxdf %>%
  filter(countyfp==191 & doy <= 59 | doy >= 335 & !is.na(tmax)) %>% #day 59= Feb 28, 335= Dec. 1
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax) #not sig

```

### Multiple regression -- Quadratic time trend

```{r quadratic temp trend}

winnewinter$yearsq <- winnewinter$year^2 #square explanatory variables for quadratic

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

#with the fitted values, create a non-linear trend
ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")

```

### Download NASS corn yield data

```{r yield download}

# set our API key with NASS
nassqs_auth(key = "B9113AF8-85C4-3CEE-8D93-6E885D49E24F") #Here put in API code from USDA QuickStats service

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

## Assignment

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?
```{r, warning=FALSE}

winnecorn <- cornyields %>%
  filter(county_ansi == "191")

cornlm <- lm(yield ~ year, data = winnecorn)
summary(cornlm) #P=1.77e-13 R^2= 0.755

ggplot(winnecorn, mapping = aes(x = year, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Year", y = "Corn Yield") +
  geom_smooth(method = lm, se=TRUE, color="#78917E", fill="#C5DDB3")

```

**Figure 1.** Linear regression of corn yields over time (years) in Winneshieck County, Iowa.

There is a significant positive correlation between corn yields and years in Winneshieck County, with an R-squared value of 0.755 and a P-value of 1.77e-13.


### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 
```{r, warning=FALSE}

winnecorn$yearsq <- winnecorn$year^2 #square explanatory variables for quadratic

lm_cornquad <- lm(yield ~ year + yearsq, winnecorn)
summary(lm_cornquad)
winnecorn$fitted <- lm_cornquad$fitted.values

#with the fitted values, create a non-linear trend
ggplot(winnecorn) +
  geom_point(mapping = aes(x = year, y = yield)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "Year", y = "Corn Yield")

```
**Figure 2.** Quadratic fit of corn yields over time (years) in Winneshieck County, Iowa.


When we fit a quadratic line to the data, we find that it follows very closely to a linear regression, suggesting a fairly linear relationship between corn yields and years in Winneshieck County. There is no evidence of slowing yield growth in the model.


### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.
```{r, warning=FALSE}

```

### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.
```{r, warning=FALSE}

```

### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.
```{r, warning=FALSE}

```

### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.
```{r, warning=FALSE}

```

### Bonus: Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.

### Bonus #2: Challenge question - map trends in corn yields by county across Iowa. Interpret your map.
